Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-08-04 tex/nusselt_paper.tex
submission versio for prfluids

\subsection{Evolved fluid numbers \& flow morphology}

@@

In 2D, low Ma flows (e.g., $\epsilon = 10^{-4}$)
display the classic narrow downflow and broad upflow lanes of stratified
convection (Fig. \ref{fig:entropy_snapshots}a).
At high Ma (e.g., $\epsilon = 0.5$, Ra/Ra$_{\text{crit}} \gtrsim 10^3$), 
bulk thermodynamic structures are similar but
shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}b\&c), as reported previously
\cite{cattaneo&all1990, malagoli&all1990}.
At large Ra, the diffusion timescale becomes long (\ref{eqn:timescales}), 
and
thermodynamic structures form small eddies which traverse the
domain repeatedly before diffusing (Fig. \ref{fig:entropy_snapshots}c).
In 3D, the same upflow/downflow asymmetry is seen, but downflow lanes no
longer span the depth of the domain (Fig. \ref{fig:entropy_snapshots}d-f).
As evidenced by the colorbar scalings, the
amplitudes of thermodynamic fluctuations scale with $\epsilon$.
\begin{figure}[t!]
\includegraphics[width=\textwidth]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved flows roughly
$\{$140, 60$\}$ $t_b$ after the start of simulations for $\{$2D, 3D$\}$ runs. 
The time- and horizontally-averaged profile is removed in vertical slices (a)-(d). 
The time- averaged mean value is removed in horizontal slices (e) and (f).
(a) A low Ma flow at moderate Ra. (b) A high Ma flow at the same Ra as in (a).
(c) A high Ma flow at high Ra. (d)-(f) A low Ma 3D run at the same parameters as in (a),
where (d) is a vertical ($x$, $z$) slice at $y = L_y/2$, (e) is a horizontal slice
at $z = L_z$, and (f) is a horizontal slice at $z = L_z/2$.
Shock systems can be seen in the upper atmosphere of the high Ma flows,
for example at $(x, z) \sim (5, 15-19)$ in (b) and $(x, z) \sim 
(50, 15-19)$ in (c).
\label{fig:entropy_snapshots} }
\end{figure}

@@

\subsection{Evolved stratification}

In the evolved state, the flows can change the density stratification.
In Fig. \ref{fig:nrho_v_ra}a, we measure the 
time- and horizontally-averaged density profile in two ways. Empty symbols
show the number of density scale heights between the maximum and minimum of the
atmospheric density profile.  Solid symbols
show the number of density scale heights between the top and bottom of the atmosphere. 
We find that near-sonic and supersonic flows support significant, 
persistent density inversions in the boundary layers, as
was reported previously \cite{brandenburg&all2005}.  This is visible when
solid symbols lie below empty symbols.  We find this in 2D and 3D, even
at very large $\epsilon$.  Sample evolved density profiles are displayed in
Fig. \ref{fig:nrho_v_ra}b.

>>>>

\vspace{-0.5cm}
\subsection{Evolved fluid numbers \& flow morphology}

@@

In 2D, low Ma flows (e.g., $\epsilon = 10^{-4}$)
display the classic narrow downflow and broad upflow lanes of stratified
convection (Fig. \ref{fig:entropy_snapshots}a).
At high Ma (e.g., $\epsilon = 0.5$, Ra/Ra$_{\text{crit}} \gtrsim 10^3$), 
bulk thermodynamic structures are similar but
shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}b\&c), as reported previously
\cite{cattaneo&all1990, malagoli&all1990}.
At large Ra, the diffusion timescale becomes long (\ref{eqn:timescales}), 
and
thermodynamic structures form small eddies which traverse the
domain repeatedly before diffusing (Fig. \ref{fig:entropy_snapshots}c).
As evidenced by the colorbar scalings, the
amplitudes of thermodynamic fluctuations scale with $\epsilon$.

In 3D, the same upflow/downflow asymmetry is seen, but other aspects 
of the flow are distinctly different.  Fig. \ref{fig:entropy_snapshots}d-f show select
snapshots of a 3D simulation with the same input parameters as the 2D case in Fig. \ref{fig:entropy_snapshots}a.
In 2D, large-scale, coherent 
spinners dominate the  flow, leading to a single upflow and downflow.  
New downflowing plumes at the upper boundary are efficiently swept into the large 
coherent structure (near $x \sim 1$ and spanning the vertical domain).  The 
behavior of downflows in 3D is strikingly different (Fig. \ref{fig:entropy_snapshots}d).
In 3D, many individual plumes detach from 
the upper boundary, but do not organize into a single dominant downflow 
in the same fashion.  Horizontal cuts near the top of the domain 
(Fig. \ref{fig:entropy_snapshots}e) reveal 
a network of narrow downflow lanes surrounding broad upflows. 
Stronger clusters of downflows near the surface are linked to 
sheets of low entropy at the midplane of the domain (Fig. \ref{fig:entropy_snapshots}f).
As the flows evolve in time, new downflows appear at the top of the domain in 
the middle of upflows and join the surrounding downflow network, 
causing the convective structures to fragment.  There is no preferred orientation 
in the newly forming downflows, and the convective flow field constantly evolves, 
which appears to prevent the occurrence of either spinners or windy states.

\begin{figure}[t!]
\includegraphics[width=\textwidth]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved flows roughly
$\{$140, 60$\}$ $t_b$ after the start of simulations for $\{$2D, 3D$\}$ runs. 
The time- and horizontally-averaged profile is removed in vertical slices (a)-(d). 
The time- averaged mean value is removed in horizontal slices (e) and (f).
(a) A low Ma flow at moderate Ra. (b) A high Ma flow at the same Ra as in (a).
(c) A high Ma flow at high Ra. 
Shock systems can be seen in the upper atmosphere of the high Ma flows,
for example at $(x, z) \sim (5, 15-19)$ in (b) and $(x, z) \sim 
(50, 15-19)$ in (c).
(d)-(f) A low Ma 3D run at the same parameters as in (a),
where (d) is a vertical ($x$, $z$) slice at $y = L_y/2$, (e) is a horizontal slice
at $z = L_z$, and (f) is a horizontal slice at $z = L_z/2$.
\label{fig:entropy_snapshots} }
\end{figure}

@@

\vspace{-0.5cm}
\subsection{Evolved stratification}

In the evolved state, the flows can change the density stratification.
In Fig. \ref{fig:nrho_v_ra}a, we measure the 
time- and horizontally-averaged density profile in two ways. Empty symbols
show the number of density scale heights between the maximum and minimum of the
atmospheric density profile.  Solid symbols
show the number of density scale heights between the top and bottom of the atmosphere. 
We find that near-sonic and supersonic flows support significant, 
persistent density inversions in the boundary layers, as
was reported previously \cite{brandenburg&all2005}.  This is visible when
solid symbols lie below empty symbols.  We find this in 2D and 3D, even
at very large $\epsilon$.

Sample evolved density profiles are displayed in Fig. \ref{fig:nrho_v_ra}b.
The natural log of the temporally and horizontally averaged density profile, 
$\ln \rho = \ln \rho_0 + \ln\rho_1$, is displayed for four cases.  At low $\epsilon$
(dotted green line), the density stratification is, to first order, unchanged from
the initial density stratification.  At high $\epsilon$, in both 2D (solid purple line)
and 3D (dashed purple line and dash-dot-dot red line), the evolved stratification differs
significantly from the initial state and does not increase monotonically with depth.
To measure the number of density scale heights between two points in the atmosphere,
$z_1$ \& $z_2$, we calculate $n_\rho(z_1, z_2) = \ln\rho(z_2) - \ln\rho(z_1)$.  Thus,
the values plotted in Fig. \ref{fig:nrho_v_ra}a for the cases in Fig. \ref{fig:nrho_v_ra}b
can be directly read off.  For example, at $\epsilon = 1$ and Ra/Ra$_{\text{crit}} \sim 10^5$
(dash-dot-dot red line), measuring the stratification between the boundaries retrieves
$n_\rho(L_z, 0) \approx -0.3$, but measuring between the maximum and minimum value of
the profile retrieves $n_\rho(\text{min, max}) \approx 1.3$.
